Measurement and Probability in Fuzzy Quantum Theories

### 2017-04-18 prop_iop.tex
prop.tex \section{Expectation Values for IVPMs} Updating...

\section{Expectation Values for IVPMs (Version 1)}

@@

There are various kind of integrals defined for IVPMs, and the most
famous one is called the Choquet integral~\cite{Choquet1954,GilboaSchmeidler1994,Grabisch2016}.

\begin{definition}Consider the following.
\begin{itemize}
\item $X=\sum_{i=1}^{N}x_{i}\mathbf{1}_{E_{i}}$ be a random variable with
$x_{1}\le x_{2}\le\ldots\le x_{N-1}\le x_{N}$ and $E_{i}\cap E_{j}=\emptyset$.
\item $\bar{\mu}:\events\rightarrow\mathscr{I}$ a classical IVPM such that
$l:\events\rightarrow\left[0,1\right]$ and $r:\events\rightarrow\left[0,1\right]$
are the left-end and the right-end of $\bar{\mu}$, respectively,
i.e., $\bar{\mu}\left(E\right)=\left[l\left(E\right),r\left(E\right)\right]$.
\item Multiplying a negative number switching the left and right end of
an interval 
\begin{equation} 
x[l,r]=\cases{ 
[xl,xr] & for $x\ge0$ ;\\
[xr,xl] & for $x\le0$ .
}
\end{equation}
\end{itemize}
Then, the Choquet integral of $X$ with respect to $\bar{\mu}$ is
\begin{equation}
\int X\rmd\bar{\mu}=\sum_{i=1}^{N}x_{i}\left[l\left(\bigcup_{j\ge i}E_{j}\right)-l\left(\bigcup_{j\ge i+1}E_{j}\right),r\left(\bigcup_{j\ge i}E_{j}\right)-r\left(\bigcup_{j\ge i+1}E_{j}\right)\right]\textrm{ .}\label{eq:interval-expectation}
\end{equation}
\end{definition}

Similar to equation~(\ref{eq:expectation}), equation~(\ref{eq:interval-expectation})
can be considered as weighted average of the possible outcomes of
$X$ as well, despite the ``weight''
\[
\left[l\left(\bigcup_{j\ge i}E_{j}\right)-l\left(\bigcup_{j\ge i+1}E_{j}\right),r\left(\bigcup_{j\ge i}E_{j}\right)-r\left(\bigcup_{j\ge i+1}E_{j}\right)\right]\subseteq\left[l\left(E_{i}\right),r\left(E_{i}\right)\right]=\bar{\mu}\left(E_{i}\right)
\]
is tighter than the interval-valued probability of the corresponding
event. Since only one event can happen at once, we don't have to be
so pessimistic to accumulate all possible fuzziness together, and
expect $\sum_{i=1}^{N}x_{i}\bar{\mu}\left(E_{i}\right)$ to happen.

@@

Consider a quantum IVPM~$\bar{\mu}$ such that $l:\events\rightarrow\left[0,1\right]$
and $r:\events\rightarrow\left[0,1\right]$ are the left-end and the
right-end of $\bar{\mu}$, respectively, i.e., $\bar{\mu}\left(E\right)=\left[l\left(E\right),r\left(E\right)\right]$.
Given the same observable with an extra condition~$x_{1}\le x_{2}\le\ldots\le x_{N-1}\le x_{N}$,
we can combine equation~(\ref{eq:interval-expectation}) and (\ref{eq:quantum-expectation})
to define the expectation value of $\mathbf{O}$ with respect to $\bar{\mu}$
as
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\sum_{i=1}^{N}\lambda_{i}\left[l\left(\sum_{j\ge i}P_{j}\right)-l\left(\sum_{j\ge i+1}P_{j}\right),r\left(\sum_{j\ge i}P_{j}\right)-r\left(\sum_{j\ge i+1}P_{j}\right)\right]\textrm{ .}\label{eq:quantum-interval-expectation}
\end{equation}
Let $\events\left(\mathbf{O}\right)$ denote the minimal sub-event
space containing $\left\{ P_{i}\right\} _{i=1}^{N}$. Then, we have
the corresponding theorem.

\begin{thm}Given a Hilbert space of $d\ge3$, for every quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
and any observable~$\mathbf{O}$, if $\bar{\mu}$ is convex, then
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\left[\min_{\rho\in\mathrm{core}\left(\bar{\mu},\events\left(\mathbf{O}\right)\right)}\expval{\mathbf{O}}_{\rho},\max_{\rho\in\mathrm{core}\left(\bar{\mu},\events\left(\mathbf{O}\right)\right)}\expval{\mathbf{O}}_{\rho}\right]\textrm{ .}\label{eq:quantum-interval-expectation-core}
\end{equation}
Conversely, if equation~(\ref{eq:quantum-interval-expectation-core})
holds for every observable, then $\bar{\mu}$ is convex.\end{thm}

\section{Expectation Values for IVPMs (Version 2)}

We may need to understand how to define expectation values of an observable
for quantum IVPMs. Recall an observable is the quantum version of
a random variable. Hence, in order to define expectation values of
an observable, we need to understand the definition of expectation
values of a random variable for an classical IVPM. Since the expectation
value of random variable~$X$ is the integral $\int X\rmd\pmeas$~\cite{Kolmogorov1950},
we need to understand the meaning of integration on IVPMs. In the
rest of the section, we will consider the real-valued and interval-valued
probability measure defined in example~\ref{ex1} and \ref{ex3},
and see how to compute the expectation value of different random variables.
To simplify the discussion later, we will only consider a random variable~$X:\Omega\rightarrow\mathbb{R}$
which is a step function, i.e., $X=\sum_{i=N^{-}}^{N^{+}}x_{i}\mathbf{1}_{E_{i}}$,
where 
\begin{itemize}
\item $x_{i}\ge0$ for $i\ge0$ and $x_{i}\le0$ for $i\le0$.
\item $E_{N^{+}}\subseteq E_{N^{+}-1}\subseteq\ldots\subseteq E_{2}\subseteq E_{1}\subseteq\Omega\backslash E_{-1}$,
$E_{N^{-}}\subseteq E_{N^{-}+1}\subseteq\ldots\subseteq E_{-2}\subseteq E_{-1}\subseteq\Omega\backslash E_{1}$,
and $\Omega=E_{0}$.
\item $\mathbf{1}_{E}$ is the indicator function defined by 
\begin{equation} 
\mathbf{1}_{A}(x)=\cases{ 
1 & if $x\in A$ ;\\ 
0 & if $x\notin A$. 
}
\end{equation}
\end{itemize}
Given a random variable~$X$ and a probability measure~$\pmeas$,
the expectation value of $X$ is 
\begin{eqnarray}
\int X\rmd\pmeas & = & \sum_{i=N^{-}}^{N^{+}}x_{i}\pmeas\left(E_{i}\right)\textrm{ .}\label{eq:expectation-1}
\end{eqnarray}

There are various kind of integrals defined for IVPMs, and the most
famous one is called the Choquet integral~\cite{Choquet1954,GilboaSchmeidler1994,Grabisch2016}.

\begin{definition}Given the random variable~$X$ as previous paragraph,
and a classical IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$.
Then, the Choquet integral of $X$ with respect to $\bar{\mu}$ is
\begin{equation}
\int X\rmd\bar{\mu}=\sum_{i=N^{-}}^{N^{+}}x_{i}\bar{\mu}\left(E_{i}\right)\textrm{ ,}\label{eq:interval-expectation-1}
\end{equation}
where multiplying a negative number switching the left and right end
of an interval 
\begin{equation} 
x[l,r]=\cases{ 
[xl,xr] & for $x\ge0$ ;\\
[xr,xl] & for $x\le0$ .
}
\end{equation}
\end{definition}

>>>>

\section{Expectation Values for IVPMs}

@@

There are various kind of integrals defined for IVPMs, and the most
famous one is called the Choquet integral~\cite{Vitali1925,Choquet1954,GilboaSchmeidler1994,Grabisch2016}.

\begin{definition}Consider the following.
\begin{itemize}
\item Let $X=\sum_{i=N^{-}}^{N^{+}}x_{i}\mathbf{1}_{E_{i}}$ be a positive
random variable with 
\[
x_{N^{-}}\le x_{N^{-}+1}\le\ldots\le x_{-2}\le x_{-1}\le x_{0}=0\le x_{1}\le x_{2}\le\ldots\le x_{N^{+}-1}\le x_{N^{+}}
\]
 and $E_{i}\cap E_{j}=\emptyset$.
\item Let $E_{i}'=\bigcup_{j\ge i}E_{j}=\set{\omega\in\Omega}{X\left(\omega\right)\ge x_{i}}\in\events$
for $i>0$, and $E_{i}'=\bigcup_{j\le i}E_{j}=\set{\omega\in\Omega}{X\left(\omega\right)\le x_{i}}\in\events$
for $i<0$.
\item $\bar{\mu}:\events\rightarrow\mathscr{I}$ a classical IVPM such that
$l:\events\rightarrow\left[0,1\right]$ and $r:\events\rightarrow\left[0,1\right]$
are the left-end and the right-end of $\bar{\mu}$, respectively,
i.e., $\bar{\mu}\left(E\right)=\left[l\left(E\right),r\left(E\right)\right]$.
\item Multiplying a negative number switching the left and right end of
an interval
\begin{equation} 
x[l,r]=\cases{ 
[xl,xr] & for $x\ge0$ ;\\
[xr,xl] & for $x\le0$ .
}
\end{equation}
\item We also need to define the sign function as 
\begin{equation} 
\mathrm{sign}=\cases{ 
1 & for $x\ge0$ ;\\
0 & for $x=0$ ;\\
-1 & for $x\le0$ .
}
\end{equation}
\end{itemize}
Then, the Choquet integral of $X$ with respect to $\bar{\mu}$ is
\begin{equation}
\int X\rmd\bar{\mu}=\sum_{i=N^{-}}^{N^{+}}x_{i}\left[l\left(E_{i}'\right)-l\left(E_{i+\mathrm{sign}\left(i\right)}'\right),r\left(E_{i}'\right)-r\left(E_{i+\mathrm{sign}\left(i\right)}'\right)\right]\textrm{ .}\label{eq:interval-expectation}
\end{equation}
\end{definition}

Similar to equation~(\ref{eq:expectation}), equation~(\ref{eq:interval-expectation})
can be considered as weighted average of the possible outcomes of
$X$ as well, despite the ``weight''
\[
\left[l\left(E_{i}'\right)-l\left(E_{i+1}'\right),r\left(E_{i}'\right)-r\left(E_{i+1}'\right)\right]\subseteq\left[l\left(E_{i}\right),r\left(E_{i}\right)\right]=\bar{\mu}\left(E_{i}\right)
\]
is tighter than the interval-valued probability of the corresponding
event. Since only one event can happen at once, we don't have to be
so pessimistic to accumulate all possible fuzziness together, and
expect $\sum_{i=1}^{N}x_{i}\bar{\mu}\left(E_{i}\right)$ to happen.

@@

Consider a quantum IVPM~$\bar{\mu}$ such that $l:\events\rightarrow\left[0,1\right]$
and $r:\events\rightarrow\left[0,1\right]$ are the left-end and the
right-end of $\bar{\mu}$, respectively, i.e., $\bar{\mu}\left(E\right)=\left[l\left(E\right),r\left(E\right)\right]$.
Given the same observable with an extra condition~
\[
x_{N^{-}}\le x_{N^{-}+1}\le\ldots\le x_{-2}\le x_{-1}\le x_{0}=0\le x_{1}\le x_{2}\le\ldots\le x_{N^{+}-1}\le x_{N^{+}}
\]
and let $P_{i}'=\sum_{j\ge i}P_{j}\in\events$ for $i>0$, and $P_{i}'=\sum_{j\le i}P_{j}\in\events$
for $i<0$, we can combine equation~(\ref{eq:interval-expectation})
and (\ref{eq:quantum-expectation}) to define the expectation value
of $\mathbf{O}$ with respect to $\bar{\mu}$ as
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\sum_{i=1}^{N}\lambda_{i}\left[l\left(P_{i}'\right)-l\left(P_{i+\mathrm{sign}\left(i\right)}'\right),r\left(P_{i}'\right)-r\left(P_{i+\mathrm{sign}\left(i\right)}'\right)\right]\textrm{ .}\label{eq:quantum-interval-expectation}
\end{equation}
Let $\events\left(\mathbf{O}\right)$ denote the minimal sub-event
space containing $\left\{ P_{i}\right\} _{i=1}^{N}$. Then, we have
the corresponding theorem.
